Analysis of Cell-Free Synthetic Circuits using Sequence Specific Constraints Based Modeling

### 2016-12-15 Paper_v1.tex
draft_v1

\usepackage{natbib}
\usepackage{setspace}
\usepackage{xkeyval}
\usepackage{url}  % Formatting web addresses
\usepackage{ifthen}  % Conditional
\usepackage{multicol}   %Columns
\usepackage[utf8]{inputenc} %unicode support
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{mathtools}
\usepackage{epsfig}
\usepackage{epstopdf}
\usepackage{graphicx}
\usepackage[margin=0.1pt,font=footnotesize,labelfont=bf]{caption}
\usepackage[version=3]{mhchem} % Formula subscripts using \ce{}
\usepackage[T1]{fontenc}       % Use modern font encodings
%\usepackage[square,sort,comma,numbers,sort&compress]{natbib}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% If issues arise when submitting your manuscript, you may want to
%% un-comment the next line.  This provides information on the
%% version of every file you have used.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%\listfiles

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% The "tocentry" environment can be used to create an entry for the
%% graphical table of contents. It is given here as some journals
%% require that it is printed as part of the abstract page. It will
%% be automatically moved as appropriate.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{tocentry}

Some journals require a graphical entry for the Table of Contents.
This should be laid out ``print ready'' so that the sizing of the
text is correct.

Inside the \texttt{tocentry} environment, the font used is Helvetica
8\,pt, as required by \emph{Journal of the American Chemical
Society}.

The surrounding frame is 9\,cm by 3.5\,cm, which is the maximum
permitted for  \emph{Journal of the American Chemical Society}
graphical table of content entries. The box will not resize if the
content is too big: instead it will overflow the edge of the box.

This box and the associated title will always be printed on a
separate page at the end of the document.

\end{tocentry}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% The abstract environment will automatically gobble the contents
%% if an abstract is not used by the target journal.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{abstract}
In this study, we used sequence specific constraints based modeling to evaluate the performance of synthetic circuits in an \emph{E.~coli} TX-TL system.
A core \emph{E.~coli} metabolic model, consisting of XX metabolites and YY reactions, was developed from literature [REF].
This model, which described glycolysis, pentose phosphate pathway, amino acid biosynthesis and degradation and energy metabolism, was then augmented with
sequence specific descriptions of genetic circuits which included mechanistic models of promoter function, transcription and translation.
Thus, unlike other synthetic biology modeling efforts, sequence specific constraints based modeling explicitly couples
the transcription and translation of circuit components with the availability of metabolic resources.
Model parameters were largely taken from literature; our approach had very few adjustable parameters thereby allowing the a first principles prediction of circuit performance.
We tested this approach by first simulating $\sigma_{70}$-induced deGFP expression and then expanded these studies to more complex multicomponent circuits.
First principles predictions of circuit performance were consistent with measurements for a variety of cases.
Further, global sensitivity analysis identified the key metabolic processes that controlled circuit performance.
Taken together, sequence specific constraints based modeling offers a novel means to \emph{a~priori} estimate the performance of cell free synthetic circuits.
\end{abstract}

@@

In this study, we used sequence specific constraints based modeling to evaluate the performance of synthetic circuits in an \emph{E.~coli} TX-TL system.
A core \emph{E.~coli} cell free metabolic model, consisting of XX metabolites and YY reactions, was developed from literature [REF].
This model, which described glycolysis, pentose phosphate pathway, amino acid biosynthesis and degradation and energy metabolism, was then augmented with
sequence specific descriptions of genetic circuits which included mechanistic models of promoter function, transcription and translation.
Thus, sequence specific constraints based modeling explicitly couples
the transcription and translation of circuit components with the availability of metabolic resources.
Model parameters were largely taken from literature; our approach had very few adjustable parameters thereby allowing the a first principles prediction of circuit performance.
We tested this approach by first simulating $\sigma_{70}$-induced deGFP expression and then expanded these studies to more complex multicomponent circuits.
First principles predictions of circuit performance were consistent with measurements for a variety of cases.
Further, global sensitivity analysis identified the key metabolic processes that controlled circuit performance.
Taken together, sequence specific constraints based modeling offers a novel means to \emph{a~priori} estimate the performance of cell free synthetic circuits.

@@

\subsection{Outline}

The document layout should follow the style of the journal concerned.
Where appropriate, sections and subsections should be added in the
normal way. If the class options are set correctly, warnings will be
given if these should not be present.

\subsection{References}

The class makes various changes to the way that references are
handled.  The class loads \textsf{natbib}, and also the
appropriate bibliography style.  References can be made using
the normal method; the citation should be placed before any
punctuation, as the class will move it if using a superscript
citation style
The use of \textsf{natbib} allows the use of the various citation
commands of that package: \citeauthor{Abernethy2003} have shown
something, in \citeyear{Cotton1999}, or as given by
Ref.~\citenum{Mena2000}.  Long lists of authors will be
automatically truncated in most article formats, but not in
supplementary information or reviews. If you
encounter problems with the citation macros, please check that
your copy of \textsf{natbib} is up to date. The demonstration
database file \texttt{achemso-demo.bib} shows how to complete
entries correctly. Notice that ``\latin{et al.}'' is auto-formatted
using the \texttt{\textbackslash latin} command.

Multiple citations to be combined into a list can be given as
a single citation.  This uses the \textsf{mcitelus} package
\cite{Johnson1972,*Arduengo1992,*Eisenstein2005,*Arduengo1994}.
Citations other than the first of the list should be indicated
with a star. If the \textsf{mcitelus} package is not installed,
the standard bibliography tools will still work but starred
references will be ignored. Individual references can be referred
to using \texttt{\textbackslash mciteSubRef}:
``ref.~\mciteSubRef{Eisenstein2005}''.

The class also handles notes to be added to the bibliography.  These
should be given in place in the document \bibnote{This is a note.
The text will be moved the the references section.  The title of the
section will change to ``Notes and References''.}.  As with
citations, the text should be placed before punctuation.  A note is
also generated if a citation has an optional note.  This assumes that
the whole work has already been cited: odd numbering will result if
this is not the case .

\subsection{Floats}

New float types are automatically set up by the class file.  The
means graphics are included as follows (Scheme~\ref{sch:example}).  As
illustrated, the float is ``here'' if possible.
\begin{scheme}
  Your scheme graphic would go here: \texttt{.eps} format\\
  for \LaTeX\, or \texttt{.pdf} (or \texttt{.png}) for pdf\LaTeX\\
  \textsc{ChemDraw} files are best saved as \texttt{.eps} files:\\
  these can be scaled without loss of quality, and can be\\
  converted to \texttt{.pdf} files easily using \texttt{eps2pdf}.\\
  %\includegraphics{graphic}
  \caption{An example scheme}
  \label{sch:example}
\end{scheme}

\begin{figure}
  As well as the standard float types \texttt{table}\\
  and \texttt{figure}, the class also recognises\\
  \texttt{scheme}, \texttt{chart} and \texttt{graph}.
  \caption{An example figure}
  \label{fgr:example}
\end{figure}

Charts, figures and schemes do not necessarily have to be labelled or
captioned.  However, tables should always have a title. It is
possible to include a number and label for a graphic without any
title, using an empty argument to the \texttt{\textbackslash caption}
macro.

The use of the different floating environments is not required, but
it is intended to make document preparation easier for authors. In
general, you should place your graphics where they make logical
sense; the production process will move them if needed.

\subsection{Math(s)}

The \textsf{achemso} class does not load any particular additional
support for mathematics.  If packages such as \textsf{amsmath} are
required, they should be loaded in the preamble.  However,
the basic \LaTeX\ math(s) input should work correctly without
this.  Some inline material \( y = mx + c \) or $ 1 + 1 = 2 $
followed by some display. \[ A = \pi r^2 \]

It is possible to label equations in the usual way (Eq.~\ref{eqn:example}).
\begin{equation}
  \frac{\mathrm{d}}{\mathrm{d}x} \, r^2 = 2r \label{eqn:example}
\end{equation}
This can also be used to have equations containing graphical
content. To align the equation number with the middle of the graphic,
rather than the bottom, a minipage may be used.
\begin{equation}
  \begin{minipage}[c]{0.80\linewidth}
    \centering
    As illustrated here, the width of \\
    the minipage needs to allow some  \\
    space for the number to fit in to.
    %\includegraphics{graphic}
  \end{minipage}
  \label{eqn:graphic}
\end{equation}

\section{Experimental}

The usual experimental details should appear here.  This could
include a table, which can be referenced as Table~\ref{tbl:example}.
Notice that the caption is positioned at the top of the table.
\begin{table}
  \caption{An example table}
  \label{tbl:example}
  \begin{tabular}{ll}
    \hline
    Header one  & Header two  \\
    \hline
    Entry one   & Entry two   \\
    Entry three & Entry four  \\
    Entry five  & Entry five  \\
    Entry seven & Entry eight \\
    \hline
  \end{tabular}
\end{table}

Adding notes to tables can be complicated.  Perhaps the easiest
method is to generate these using the basic
\texttt{\textbackslash textsuperscript} and
\texttt{\textbackslash emph} macros, as illustrated (Table~\ref{tbl:notes}).
\begin{table}
  \caption{A table with notes}
  \label{tbl:notes}
  \begin{tabular}{ll}
    \hline
    Header one                            & Header two \\
    \hline
    Entry one\textsuperscript{\emph{a}}   & Entry two  \\
    Entry three\textsuperscript{\emph{b}} & Entry four \\
    \hline
  \end{tabular}

  \textsuperscript{\emph{a}} Some text;
  \textsuperscript{\emph{b}} Some more text.
\end{table}

The example file also loads the optional \textsf{mhchem} package, so
that formulas are easy to input: \texttt{\textbackslash ce\{H2SO4\}}
gives \ce{H2SO4}.  See the use in the bibliography file (when using
titles in the references section).

The use of new commands should be limited to simple things which will
not interfere with the production process.  For example,
\texttt{\textbackslash mycommand} has been defined in this example,
to give italic, mono-spaced text: \mycommand{some text}.

\section{Extra information when writing JACS Communications}

When producing communications for \emph{J.~Am.\ Chem.\ Soc.}, the
class will automatically lay the text out in the style of the
journal. This gives a guide to the length of text that can be
accommodated in such a publication. There are some points to bear in
mind when preparing a JACS Communication in this way.  The layout
produced here is a \emph{model} for the published result, and the
outcome should be taken as a \emph{guide} to the final length. The
spacing and sizing of graphical content is an area where there is
some flexibility in the process.  You should not worry about the
space before and after graphics, which is set to give a guide to the
published size. This is very dependant on the final published layout.

You should be able to use the same source to produce a JACS
Communication and a normal article.  For example, this demonstration
file will work with both \texttt{type=article} and
\texttt{type=communication}. Sections and any abstract are
automatically ignored, although you will get warnings to this effect.

>>>>

\usepackage{natbib}
\usepackage{setspace}
\usepackage{xkeyval}
\usepackage{url}  % Formatting web addresses
\usepackage{ifthen}  % Conditional
\usepackage{multicol}   %Columns
\usepackage[utf8]{inputenc} %unicode support
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{mathtools}
\usepackage{epsfig}
\usepackage{epstopdf}
\usepackage{graphicx}
\usepackage{textcomp}
\usepackage{multirow}
\usepackage{booktabs}
\usepackage[margin=0.1pt,font=footnotesize,labelfont=bf]{caption}
\usepackage[version=3]{mhchem} % Formula subscripts using \ce{}
\usepackage[T1]{fontenc}       % Use modern font encodings
%\usepackage[square,sort,comma,numbers,sort&compress]{natbib}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% If issues arise when submitting your manuscript, you may want to
%% un-comment the next line.  This provides information on the
%% version of every file you have used.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%\listfiles

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% The "tocentry" environment can be used to create an entry for the
%% graphical table of contents. It is given here as some journals
%% require that it is printed as part of the abstract page. It will
%% be automatically moved as appropriate.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% The abstract environment will automatically gobble the contents
%% if an abstract is not used by the target journal.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{abstract}
In this study, we used sequence specific constraints based modeling to evaluate the performance of synthetic circuits in an \emph{E.~coli} cell-free protein synthesis system.
A core \emph{E.~coli} metabolic model, consisting of glycolysis, pentose phosphate pathway, amino acid biosynthesis and degradation and energy metabolism, was then augmented with sequence specific descriptions of genetic circuits which included mechanistic models of promoter function, transcription and translation.
Thus, unlike other synthetic biology modeling efforts, sequence specific constraints based modeling explicitly couples the transcription and translation of circuit components with the availability of metabolic resources.
Model parameters for transcription and translation were taken from literature, allowing a first principles prediction of circuit performance.
We tested this approach by first simulating T7 induced CAT production and $\sigma_{70}$-induced deGFP expression; we then expanded these studies for a range of different proteins
First principles predictions of circuit performance were consistent with measurements for a variety of cases.
Further, global sensitivity analysis identified the key metabolic processes that controlled circuit performance in terms of productivity and yield.
A sufficient energy supply with oxidative phosphorylation is instrumental for high carbon yields, in addition the translation rate is the next step to optimize for higher productivity and yield. 
Taken together, sequence specific constraints based modeling offers a novel means to \emph{a~priori} estimate the performance of cell-free synthetic circuits.
\end{abstract}

@@

In this study, we used sequence specific constraints based modeling to evaluate the performance of synthetic circuits in an \emph{E.~coli} cell-free protein synthesis (CFPS) system.
A core \emph{E.~coli} cell free metabolic model, consisting of 132 metabolites and 281 reactions, was developed from literature \cite{Feist:2007aa}.
This model, which described glycolysis, pentose phosphate pathway, amino acid biosynthesis and degradation and energy metabolism, was then augmented with
sequence specific descriptions of genetic circuits which included mechanistic models of promoter function, transcription and translation.
Thus, sequence specific constraints based modeling explicitly couples the transcription and translation of circuit components with the availability of metabolic resources.
We tested this approach by first simulating $\sigma_{70}$-induced deGFP expression at different plasmid concentrations and then expanded this for the production of CAT and deGFP under two different promoters.
First principles predictions of circuit performance were consistent with measurements.
We then investigated the productivity and carbon yield for a range of 8 different proteins for a variety of cases. 
From this, higher carbon number proteins typically had lower productivity rates and carbon yields than that of the lower carbon number proteins. 
Further, global sensitivity analysis identified the key metabolic processes that controlled circuit performance, showing oxidative phosphorylation is instrumental for maintaining a high carbon yield and the translation rate for productivity. 
Taken together, sequence specific constraints based modeling offers a novel means to \emph{a~priori} estimate the performance of cell free synthetic circuits.

@@

\section*{Results}
\subsection{Model Derivation}
The goal of this work was first to construct a modeling framework to describe cell-free protein synthesis systems and to examine its performance in productivity and yield for a protein of interest.
One mathematical framework that has found wide use in modeling metabolism are constraint based models such as flux balance analysis (FBA).
FBA can predict how cells utilize nutrients to produce products by using the biochemical stoichiometry and thermodynamical feasibility under pseudo steady-state conditions.
Traditionally, FBA is used to model \textit{in vivo} processes, however cell-free systems do not have growth associated reactions or transport through the cell membrane.
Thus, to model cell-free metabolism we constructed a cell-free stoichiometric network by removing growth associated reactions from the MG1655 reconstruction \cite{Feist:2007aa}, and incorporating amino acid synthesis and transcription/translation associated reactions \cite{Allen:2003aa} for a protein of interest to be expressed. 
The network consisted of 281 reactions and 132 species. 
We developed a cell-free sequence specific flux balance analysis (ssFBA) with a detailed promoter model \cite{Moon:2012aa} to examine the performance of CFPS. 
We first validated the ssFBA approach by comparing simulated and measured concentrations of two proteins from two different cell-free \textit{E. coli} extract systems.
The first protein, chloramphenicol acetyltransferase (CAT), was produced under a T7 promoter in a glucose/NMP cell-free system \cite{2005_calhoun_BiotechnologyProgress} for a duration of 1 hour under glucose consumption (Fig \ref{fig:CAT_GFP_prod}A).
The second protein, dual emission green fluorescent protein (deGFP), was produced under a P70 promoter in TXTL 2.0 \textit{E. coli} extract for a duration of 8 hours under maltose consumption(Fig \ref{fig:CAT_GFP_prod}B).
The ssFBA simulations predicted the production of both proteins for the duration of both CFPS batch reactions.
Uncertainty in experimental factors such as RNA polymerase, ribosome concentrations, elongation rates or the upper bounds for oxygen and glucose consumption rates did not alter the qualitative performance of the model. 
Thus, the metabolic network and molecular description of trancription and translation were consistent with experimental measurements.
\begin{figure}[t!]
\centering
\includegraphics[width=1.00\textwidth]{./Figures/CAT_GFP_prod.pdf}
\caption{Sequence specific flux balance analysis of protein production vesus time. A. CAT production under a T7 promoter in CFPS \textit{E. coli} extract for 1 hr under glucose consumption. B. deGFP production under a P70 promoter in TXTL 2.0 \textit{E. coli} extract for 8 hr under glucose consumption. 95\% CI (blue region) over the ensemble of 100 sets.}
\label{fig:CAT_GFP_prod}
\end{figure}

Next, ssFBA predicted deGFP production as a function of plasmid concentration (Fig \ref{fig:GFP_plasmid}).
Concentration of deGFP at each plasmid concentration was calculated by multiplying the flux of deGFP synthesis by the active time of production, approximately 8 hours in TXTL 2.0 \cite{2016_Garamella_etal_TXTL}. 
The mean of the ensemble shows a good prediction against the measured deGFP levels, even though it under predicted deGFP concentration at the saturating point of 5 nM of plasmid concentration.  
However, the ensemble and the mean of the ensemble captured the overall saturating dynamics of deGFP production as a function of plasmid concentration.
\begin{figure}[t!]
\centering
\includegraphics[width=0.6\textwidth]{./Figures/GFP_plasmid.pdf}
\caption{Predicted deGFP concentration (black line) at different plasmid concentrations versus measurements of deGFP (dots) synthesized in TXTL 2.0. 95\% CI (blue region) over the ensemble of 100 sets.}
\label{fig:GFP_plasmid}
\end{figure}

These results validated our mathematical framework to model CFPS systems and predict the production of two proteins with no adjustable parameters. 
It also showed that the sequence specific reactions were sufficient to predict the production of two different proteins under different promoters and cell-free systems. 
Since the model accurately predicted the behavior of protein production, we wanted to use our mathematical framework to help us understand the performance limits of CFPS and how these shortcomings could be addressed.

\subsection{Examining CFPS Performance: Productivity}
Our next goal was to examine the performance of CFPS productivity for eight different proteins under three different cases. Each of the proteins were produced under a P70 promoter, except for CAT which was produced under a T7 promoter. In all our cases, CFPS is supplied with glucose. The first case, CFPS is supplied with amino acids and the system can synthesis amino acids (control). In the second case, CFPS is supplied with amino acids, however the system cannot synthesis amino acids (AA uptake w/o synthesis). We turned off these synthesis reactions because during the cell-free extract preparation the cells are often supplied with amino acids, thus the enzymes responsible for amino acid synthesis would not be present. In the third case, CFPS is not supplied with amino acids, but the system can synthesis them (AA synthesis w/o uptakte).    
We used ssFBA to estimate the productivity of eight proteins for each case (Fig \ref{fig:Prod_POI}A).
The second case (without amino acid synthesis) showed the highest productivity for each of the proteins, however the control case had very similar performance.
This shows the system has sufficient substrates to power the system and synthesis each protein. 
The third case had the lowest productivity for each protein, thus the addition of amino acids to CFPS extract is important for maintaining a relatively high productivity.
The qualitative trend of productivity for the three different cases was the same, however some proteins had higher productivity than others. 
For instance, in the second case FGF21 had a productivity of 17 ($\mu$M/h) whereas FII had a productivity of 3 ($\mu$M/h).
To examine this further, the mean productivity was plotted against the carbon number of each protein (Fig \ref{fig:Prod_POI}B).
The proteins with the highest productivity had the lowest carbon number whereas proteins with low productivity had higher carbon numbers.     
This inverse qualitative trend was due to the fact that larger proteins require more amino acids and substrates to assemble them resulting in lower productivity. 
\begin{figure}[t!]
\centering
\includegraphics[width=1.00\textwidth]{./Figures/Prod_POI.pdf}
\caption{CFPS productivity performace for eight proteins for the control (black), AA uptake w/o synthesis (off white), and AA synthesis w/o uptake (grey). A. Productivity normalized to the specific glucose uptake rate (Error bars represent the 95\% CI of the ensemble). B. Mean productivity versus the carbon number for each corresponding protein.}
\label{fig:Prod_POI}
\end{figure}

\subsection{Examining CFPS Performance: Carbon Yield}
Following the same outline as in examining the productivity, we calculated the carbon yield for each protein (Fig \ref{fig:Yield_POI}A). 
The same trends followed, where the case without amino acid synthesis showed the highest carbon yield and the control case had comparable performance. 
The third case (with no amino acid uptake) had the lowest yields; this is most likely because glucose is now utilized to synthesize the necessary amino acids for each protein as well as power the system. 
We also investigated the carbon contribution of glucose and each amino acid toward the production of deGFP for all three cases (Table \ref{tbl:yield_breakdown}).
For the first case, the system is relying on a mixture of glucose and some amino acids for the productin of deGFP with a carbon yield of 20.5\%.
Once amino acid synthesis is removed from the network (2nd case), each amino acid is utilized and the carbon yield is increased to 22.2\%.
Only the necessary amount of amino acids are used for the production of CAT, thus it may be hypothesized that all the glucose is used to power CFPS and does not contribute to the carbon yield. 
If that is the case, the carbon yield without glucose contribution is 100\%, whereas for the first case its 92.5\%.
Finally for the third case (without amino acid supplemented), the carbon yield reduces to 14.8\% and the system uses about double the amount of glucose as the first two cases. 
For this case glucose is being used to synthesize amino acids and power the system and this trend was seen across all proteins except for CAT.

\begin{table}[t!]
\centering
    \caption{Carbon contribution from glucose and each amino acid toward deGFP production for three cases: control, amino acid uptake without synthesis, and amino acid synthesis without uptake.}
    \renewcommand{\arraystretch}{0.95}
    \begin{tabular}{lccc} \toprule
        \parbox{3.2 cm}{Carbon \phantom{abcdefghi} Produced (mM)} & Control & \parbox{2.5 cm}{AA uptake \phantom{ai} w/o synthesis} & \parbox{2.5 cm}{AA synthesis w/o uptake} \\ \midrule
        deGFP & 9.8 & 9.6 & 9.9  \\ \toprule
        \parbox{3.2 cm}{Carbon \phantom{abcdefghi} Consumed (mM)} & \\ \midrule
        GLC & 37.3 & 33.7 & 66.9 \\ 
        ALA & 0.0 & 0.2 & -  \\ 
        ARG & 0.3 & 0.3 & - \\ 
        ASN & 0.5 & 0.4 & -  \\ 
        ASP & 0.4 & 0.6 & -  \\ 
        CYS & 0.2 & 0.1 & -  \\ 
        GLU & 2.1 & 0.6 & -  \\ 
        GLN & 0.4 & 0.3 & - \\ 
        GLY & 0.3 & 0.3 & -  \\
        HIS & 0.5 & 0.5 & - \\
        ILE & 0.0 & 0.6 & -  \\ 
        LEU & 1.0 & 1.0 & - \\ 
        LYS & 1.0 & 0.9 & - \\ 
		MET & 0.0 & 0.2 & -  \\ 
        PHE & 1.0 & 0.9 & - \\ 
        PRO & 0.4 & 0.4 & -  \\ 
        SER & 0.0 & 0.2 & - \\ 
        THR & 1.0 & 0.5 & -  \\
        TRP & 0.1 & 0.1 & -  \\ 
        TYR & 0.8 & 0.8 & - \\ 
        VAL & 0.6 & 0.7 & - \\ \hline
        Sum & 47.9 & 43.3 & 66.9  \\ \toprule
        Yield & 20.5\% & 22.2\% & 14.8\% \\ 
	Yield w/o GLC & 92.5\% & 100\% & - \\ \bottomrule
    \end{tabular}
\label{tbl:yield_breakdown}
\end{table}

Interestingly, CAT carbon yield showed similar performance for all three cases. 
Each protein has different energy requirements for transcription and translation depending on its sequence.
Thus, for the case of CAT, energy requirements for transcription or translation are low enough that it's carbon yield is not hindered in the third case.
We next investigated the effect of the carbon number of each protein to the carbon yield (Fig \ref{fig:Yield_POI}B). 
The same inverse qualitative trend is observed, however it is less significant. 
Sequence specific flux balance analysis assumes a psuedo steady state, thus intermediate metabolites cannot be accumulated within the cell-free extract.
In addition, ssFBA is solved by setting the production of the protein as the objective function.
Therefore, carbon flux will travel throughout the network to optimize the flux through the protein synthesis reaction. 
\begin{figure}[t!]
\centering
\includegraphics[width=1.00\textwidth]{./Figures/Yield_POI.pdf}
\caption{CFPS carbon yield performace for eight proteins for the control (black), AA uptake w/o synthesis (off white), and AA synthesis w/o uptake (grey). A. Carbon yield (Error bars represent the 95\% CI of the ensemble). B. Mean carbon yield versus the carbon number for each corresponding protein.}
\label{fig:Yield_POI}
\end{figure}

\subsection{Sensitivity analysis on CFPS Performance}
To better understand the effect of substrate utilization and transcription/translation parameters on CFPS performance we performed global sensitivity analysis on productivity and carbon yield for deGFP, a representative protein (Fig \ref{fig:SI_GFP}).
In examining productivity performance (Fig \ref{fig:SI_GFP}A), the significance of transcription/translation parameters were fairly constant across all three cases, with the elongation rate by ribosomes being the most sensitive.
This shows the rate of translation is instrumental in achieving high productivity, and should be the first step to investigate in order to optimize productivity, prior to examining transcription rates, this is consistent with literature findings.
Underwood and coworkers have also shown that an increase in ribosome levels does not significantly increase protein yields or rates; however, adding elongation factors increased yields by 23\% at 30 minutes \cite{2005_underwood_biotech}.
In addition, Li et al. have increased productivity of firefly lucifease by 5-fold in CFPS systems by adding and adjusting factors that affect transcription and translation such as elongation factors, ribosome recyclcing factor, release factors, chaperones, BSA, and tRNAs \cite{2014_li_PlosOne}.
In examining the substrate utilization, glucose uptake was not very important for productivity in the first two cases, but its significance increased when amino acids was removed from CFPS.
This makes sense, as amino acid synthesis can only come from glucose and becomes the only way to power protein synthesis.
On the other hand, when amino acid synthesis is removed, amino acid uptake becomes more important, as it becomes the only source of amino acids.
\begin{figure}[t!]
\includegraphics[width=1.00\textwidth]{./Figures/SI_GFP.pdf}
\caption{Total order sensitivity of deGFP productivity (A) and carbon yield (B) to specific uptake rates and transcription/translation parameters for three cases: control (black), amino acid uptake without synthesis (off white), and amino acid synthesis without uptake (gray). (Error bars represent 95\% CI of the ensemble).}
\label{fig:SI_GFP}
\end{figure}

When considering carbon yield performance (Fig \ref{fig:SI_GFP}B), the substrate and oxygen uptake fluxes become much more important while the sensitivity to the transcription/translation parameters decreases slightly.
The transcription/translation parameters had the same trend as in the productivity trend, where the ribosome elongation rate was the most sensitive compared to the transcription/translation parameters and showed significance across all parameters. 
Thus, in investigating CFPS performance in terms of productivity and yield, the ribosome elongation rate seems to be an important parameter to optimize, as been already shown in literature \cite{2005_underwood_biotech, 2014_li_PlosOne}. 
Glucose and oxygen uptake shown significant importance for carbon yield since it determines the mechanism of energy generation, which is critical to efficient protein production.
Meanwhile, productivity is determined primarily by the rate of the most downstream steps, transcription and translation.
Glucose is required for both amino acid synthesis and efficient energy generation, both of which are important for a high yield.
There is a tradeoff between energy generation to power transcription/translation as well as synthesis of amino acids which are required to assemble the protein of interest. 
Across all three cases, substrate utilization (amino acid uptake and glucose) showed to be the next important as these substrates are contributing to the carbon yield of deGFP.
Interestingly, in the first two cases, oxygen uptake has a significant effect on carbon yield.
This is most likely due to the importance of oxidative phosphorylation for efficient energy generation.
However, in the third case (with no amino acid uptake), oxygen uptake becomes insignificant compared to the other parameters, while glucose uptake was the most siginificant.
Jewett and coworkers have reported that oxidative phosphorylation still operates in cell free and also saw a decrease in yield, ranging from 1.5-fold to 4-fold, when knocking out oxidative phosphorylation reactions\cite{Jewett:2008aa}.
However, it is unknown how active oxidative phosphorylation is compared to that of \textit{in vivo} systems, thus to investigate this further we compared deGFP carbon yield to oxidative phosphorylation flux (Fig \ref{fig:oxphos_yield}).
Interestingly, oxidative phosphorylation has a somewhat linear correlation with carbon yield for the first two cases (R\textsuperscript{2}=0.77 for the control, R\textsuperscript{2}=0.79 for the second case), whereas for the third case, this same effect was not as prevalent (R\textsuperscript{2}=0.46). 
\begin{figure}[t!]
\includegraphics[width=0.7\textwidth]{./Figures/OxPhos_yield.pdf}
\caption{An ensemble of 1000 ssFBA solutions for deGFP carbon yield versus oxidative phosphorylation flux for three cases: control (black), amino acid uptake without synthesis (off white), and amino acid synthesis without uptake (gray). }
\label{fig:oxphos}
\end{figure}

When the system can utilize glucose for energy generation, the increase in oxidative phosphorylation has a positive effect on carbon yield. 
However, when glucose is used to synthesize amino acids and power CFPS, there is a limit of how much oxidative phosphorylation can effect carbon yield.
Thus, it is interesting to see how much of the energetic needs of the system are met by oxidative phosphorylation and where the remaining energy is coming from. 
To investigate the differences in productivity and carbon yields, we compared the flux distributions for deGFP predicted by ssFBA simulations for the three different cases (Fig.\ref{fig:flux}).
All cases used the first step in the pentose phosphate pathway to generate NADPH;
the carbon flux then continued through the Entner–Doudoroff pathway toward pyruvate.
The majority of the flux was distributed between lactate and acetate, representing an inefficiency in energy generation as well as wasteful byproducts.
For the first two cases, the energy source was primarily oxidative phosphorylation, and to a lesser extent the TCA cycle.
The second case with no amino acid synthesis showed to be the most efficient process (highest carbon yield).
All the NADPH produced is interconverted to NADH to power oxidative phosphorylation.
Oxidative phosphorylation lead to a high redox ratio contributing to the accumulation of acetate overflow 
However, diverting the flux away from acetate accumulation may allow more resources towards the protein of interest.
Interestingly, for the first case and more for the third case, there is a mixture of aerobic and anearobic energy generation. 
Fumerate dehydrogenase typically operates in anaerobic conditions and is less efficient than oxidative phosphorylation. 
Cell-free systems do not have a sense of regulation as \textit{in vivo} systems thus any enzyme present in the mixture will carry out its corresponding reaction.
In the third case (third row), there is a low flux through oxidative phosphorylation and high flux through fumerate dehydrogenase (fumerate to succinate). 
One of the reasons for this inefficiency is some of the NADH is interconverted to NADPH which is required for amino acid synthesis, thus the lack of NADH cannot power oxidative phosphorylation as effectively as in the first two cases.
It is currently unknown how much oxidative phosphorylation is present in cell-free systems, however our model and the sensitivity analysis suggests that efficient energy generation is essential to high carbon yields. 
Thus, removing anaerobic enzymes during the cell-free extract preparation is a potential area for improvement for CFPS performance and protein yield. 
\begin{figure}[h!]
\includegraphics[width=0.8\textwidth]{./Figures/flux.pdf}
\caption{deGFP production flux profile for glycolysis, pentose phosphate pathway, Entner-Doudoroff pathway, TCA cycle, NADPH/NADH transfer, and oxidative phosphorylation. Flux (mean $\pm$ standard deviation) across ensemble normalized to glucose uptake flux. Flux distribution for three different cases: control (top row), amino acid uptake without synthesis (second row), and amino acid synthesis without uptake (third row). Bold metabolites are allowed to accumulate; grey arrows lead to amino acid synthesis.}
\label{fig:flux}
\end{figure}

All proteins followed a similar trend where for the first two cases (when amino acids were supplied) there was a relatively high yield than compared to the third case, with the exception of CAT. 
Thus, we also performed a sensitivity analysis on CAT (Fig \ref{fig:SI_CAT}), which did not show the same trend in yield across the three cases as the other proteins.
For CAT, glucose uptake was as important to productivity as the kinetic parameters.
The oxygen uptake was the most important factor in yield and maintained this across all three cases.
The energy requirements for transcription of CAT is lower than the other proteins, thus it can accomodate its resources towards efficient energy generation.
Thus, the carbon yield is not hindered in the third case, since the system maintains the necessary amino acid synthesis rates and the required energy species.
Apart from these differences, there was a general trend of transcription and translation holding more significance for productivity and substrate uptake holding more significance for yield.

Taken together, we used sequence specific constraints based modeling to evaluate the performance of synthetic circuits in an \emph{E.~coli} cell-free protein synthesis (CFPS) system.
We have shown sequence specific constraints based modeling offers a novel means to \emph{a~priori} estimate of protein production for deGFP and CAT under two different proteins.
This modeling approach validated general trends seen in literature for productivity and yield as a function of carbon number. 
Further, global sensitivity analysis identified oxidative phosphorylation is instrumental for maintaining a high carbon yield and the translation rate is important for productivity.
The model also identified that cell-free systems can operate with aerobic and anaerobic enzymes which leads to inefficient CFPS systems and should be addressed to optimize carbon yield as seen in the globol sensitivity analysis. 
Taken together, sequence specific constraints based modeling offers a novel means to \emph{a~priori} estimate the performance of cell free synthetic circuits.
